---
title: "Day 5 Comp Session 1"
format:
  html:
    code-fold: true
    code-summary: "Show me the code!"
editor: visual
execute: 
  echo: true
  warning: false
---

#make a change

#echo controls if whether or not the r code shows up.

```{r}
library(tidyverse)
```

## Explore the starwars dataset

look at the first six lines, find the dimensions of starwars, make a histogram of character heights, and make a scatter plot of height vs. weight

```{r}

head(starwars)

dim(starwars)

tail(starwars)

#making a histogram of character heights, only needs one variable, counts of observation are presented as the bar height

ggplot(data= starwars, aes(x=height))+
  geom_histogram(color = "blue", fill= "black")
#  pretty long lower tail, could be negatively skewed


#making a scatter plot of height vs. weight
ggplot(data= starwars, aes(x=height, y= mass))+
  geom_point(aes(color=species))
```

#### Finding summary statistics

The operator you use to call an operation on a specific column, you use the \$ sign! Its accessing all the values in the column as a vector.

NA indicates a missing value\*

Use mean() function to find means.

Use sd() function to find standard deviations. (when things are symmetrical, a better representation for a data spread)

```{r}
mean_height <- mean(starwars$height, na.rm= TRUE)

mean_height

#na.rm can be set to true or false whether the na values should be removed before finding the mean value, setting na.rm removes it, so you set it to true.  
```

```{r}
# find and store the standard deviation of character heights

sd_height <- sd(starwars$height, na.rm=TRUE)
sd_height

# this is a hollistic observation from how far the values tend to be away from mean height
```

## In-line reporting using code outputs

Typing values out are not an automated way for values to exist in a report.

The mean height of star wars characters is `r mean_height` cm with a standard deviation of `r sd_height` cm.
